First and foremost, I would like to show my deepest gratitude towards my  Prof. C.V. Jawahar. I have been extremely fortunate to have him as my adviser during my course work. His constant guidance, support and motivating words have helped me become a better researcher and a better person.

I would also like to thank all the faculty members at \textsc{CVIT}, my adviser Prof. C.V. Jawahar as well as Prof. P.J. Narayanan, Prof. Jayanthi Sivaswamy, Dr. Anooop Namboodiri, Dr. Avinash Sharma and Dr. Vineet Gandhi for creating a wonderful research environment in \textsc{CVIT} to research and grow. I would also like to thank the \textsc{CVIT} administration staff, Sathya Sir, Phani Sir, Rajan Garu, Nandini and Prathima ma'am for helping me on numerous occasions.

A special thanks to two of my favorite PhD scholars, Anand Mishra and Praveen Krishnan, who have always took out time from their busy schedule to help, support and guide me. My life at \textsc{IIIT} would not be complete without the mention of my dear friends Devendra, Viresh, Mohak, Aniket, Suriya, Sanchit and Udit. I am thankful to you guys for giving me a ton of happy memories during my stay at IIIT. I would also like to thank Yashaswi, Saurabh D, Varun, Minesh, Jobin, Vinitha, Pritish, Saurabh S, Rajvi, Arunava, Koustav G, Koustav M, Aditya, Nataraj, Arpit, Sumit, Vignesh, Vijay for contributing to my work directly or indirectly. A special mention to Dr. Gaurav Harit at Indian Institute of Technology, Jodhpur who encouraged me to go for higher studies.

To my parents, my sister and Devanshu, thanks for being there and supporting me in all my ventures, successful or failures.